---
title: "Format tests"
author: "Agustin Gallo"
date: "22/5/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Method
I will divide the structure of this project in 4 sections in this section I will describe each part and in the next chapter (Results) I will show the results obtained for every section described here. So here you might expect to see the code used in the project and the description of the logic behind each section. The 4 sections in which this chapter is divided are the following.

1. Get movie and rating data. Download and create movies and rating datasets, join dataframes and create test and validation dataframes. 
2. First dummy attempt. Evaluation Criteria(RMSE) and first exploration of a dummy prediction.
3. Data exploration. Explore dataframe data to create a model hypothesis
4. Model implementation. Based on data exploration create a prediction model approach and final result
 

```{r cars, include=TRUE, echo=TRUE, warning=FALSE, results='hide'}
a <- 25
b <- 30
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
